---
title: "Report Exercice 9"
author: "Thaddäus Braun"
date: "2023-05-08"
output: html_document
---

The topic of this exercice is a comparison of the a linear regression and KNN models. In the following chunks, I am going to split the data in a test and training data set to evaluate the model performances. Before I am able to perform this step I have to pre-process the data. 
In a first step, I load the required packages and read the data. Additionally, data wrangling is performed.

```{r}
library(dplyr)
library(readr)
library(lubridate)


daily_fluxes <- read_csv("FLX_CH-Dav_FLUXNET2015_FULLSET_DD_1997-2014_1-3.csv") |>  
  
  # select only the variables we are interested in
  dplyr::select(TIMESTAMP,
                GPP_NT_VUT_REF,    # the target
                ends_with("_QC"),  # quality control info
                ends_with("_F"),   # includes all all meteorological covariates
                -contains("JSB")   # weird useless variable
                ) |>

  # convert to a nice date object
  dplyr::mutate(TIMESTAMP = ymd(TIMESTAMP)) |>

  # set all -9999 to NA
  mutate(across(where(is.numeric), ~na_if(., -9999))) |>
  
  # retain only data based on >=80% good-quality measurements
  # overwrite bad data with NA (not dropping rows)
  dplyr::mutate(GPP_NT_VUT_REF = ifelse(NEE_VUT_REF_QC < 0.8, NA, GPP_NT_VUT_REF),
                TA_F           = ifelse(TA_F_QC        < 0.8, NA, TA_F),
                SW_IN_F        = ifelse(SW_IN_F_QC     < 0.8, NA, SW_IN_F),
                LW_IN_F        = ifelse(LW_IN_F_QC     < 0.8, NA, LW_IN_F),
                VPD_F          = ifelse(VPD_F_QC       < 0.8, NA, VPD_F),
                PA_F           = ifelse(PA_F_QC        < 0.8, NA, PA_F),
                P_F            = ifelse(P_F_QC         < 0.8, NA, P_F),
                WS_F           = ifelse(WS_F_QC        < 0.8, NA, WS_F)) |> 

  # drop QC variables (no longer needed)
  dplyr::select(-ends_with("_QC"))

```

In order to specify to specify the target and predictor variables, I use the formula notation.

```{r}
#formula notation

lm(GPP_NT_VUT_REF ~ SW_IN_F + VPD_F + TA_F, data = daily_fluxes)


```

In this chunk I use the function "train" from the package "caret" to train the model. Here, I speficy the linear regression model. 

```{r}

library(caret)
library(tidyverse)

caret::train(
  form = GPP_NT_VUT_REF ~ SW_IN_F + VPD_F + TA_F, 
  data = daily_fluxes |> drop_na(),  # drop missing values
  trControl = caret::trainControl(method = "none"),  # no resampling
  method = "lm"
)


```


In this chunk I specify the knn model. Again with the train function. 

```{r}

library(caret)

caret::train(
  form = GPP_NT_VUT_REF ~ SW_IN_F + VPD_F + TA_F, 
  data = daily_fluxes |> drop_na(), 
  trControl = caret::trainControl(method = "none"),
  method = "knn"
)


```

Once the two models have been specified, the data is now split into a test and a train data set. To effectively evaluate the generalization error of a model, a crucial procedure involves segregating a portion of the data during training and keeping it completely separate and untouched for testing purposes. In this case, 70% of the data is used for training. This is a typical split to ensure that there is enough data for training that the test results are robust and also able to find a relationship. 
In a second step the distribution of of values in the training and test set is plotted.

```{r}

set.seed(123)  # for reproducibility
split <- rsample::initial_split(daily_fluxes, prop = 0.7, strata = "VPD_F")
daily_fluxes_train <- rsample::training(split)
daily_fluxes_test <- rsample::testing(split)

plot_data <- daily_fluxes_train |> 
  dplyr::mutate(split = "train") |> 
  dplyr::bind_rows(daily_fluxes_test |> 
  dplyr::mutate(split = "test")) |> 
  tidyr::pivot_longer(cols = 2:9, names_to = "variable", values_to = "value")

plot_data |> 
  ggplot(aes(x = value, y = ..density.., color = split)) +
  geom_density() +
  facet_wrap(~variable, scales = "free")



```


The purpose of data pre-processing is twofold: first, it involves preparing the data for a particular model fitting process, and second, it aims to enhance the efficacy of model training.
The {recipes} package offers a more robust approach for simultaneously specifying the formula and pre-processing steps. It seamlessly integrates with the {caret} package's train() function. To illustrate, considering the same formula mentioned earlier, and assuming we want to normalize (center and scale) the data in daily_fluxes_train, we can define a "recipe" using the pipe operator as follows:

```{r}

library(recipes)

pp <- recipes::recipe(GPP_NT_VUT_REF ~ SW_IN_F + VPD_F + TA_F, data = daily_fluxes_train) |> 
  recipes::step_center(all_numeric(), -all_outcomes()) |>
  recipes::step_scale(all_numeric(), -all_outcomes())

caret::train(
  pp, 
  data = daily_fluxes_train, 
  method = "knn",
  trControl = caret::trainControl(method = "none")
)


```

The following chunk aims to standardize the values of the data. Some algorithms explicitly necessitate the standardization of data to ensure that the values of all predictors fall within a comparable range.

```{r}
daily_fluxes |> 
  summarise(across(where(is.numeric), ~quantile(.x, probs = c(0, 0.25, 0.5, 0.75, 1), na.rm = TRUE))) |> 
  t() |> 
  as_tibble(rownames = "variable") |> 
  setNames(c("variable", "min", "q25", "q50", "q75", "max"))


```

Furthermore, there are more steps required to actually transform "daily_fluxes_train". 

```{r}

pp_prep <- recipes::prep(pp, training = daily_fluxes_train) 

daily_fluxes_juiced <- recipes::juice(pp_prep)

daily_fluxes_baked <- recipes::bake(pp_prep, new_data = daily_fluxes_train)

# confirm that juice and bake return identical objects when given the same data
all_equal(daily_fluxes_juiced, daily_fluxes_baked)

daily_fluxes_baked <- recipes::bake(pp_prep, new_data = daily_fluxes_train)

# confirm that juice and bake return identical objects when given the same data
all_equal(daily_fluxes_juiced, daily_fluxes_baked)

```

This part of the code replaces missing values with the best guess. This step is called imputation. Here, I use KNN with five neighbors.

```{r}
pp |> 
  step_impute_median(all_predictors())
pp |> 
  step_impute_knn(all_predictors(), neighbors = 5)

```

Next, I fit the linear regression model "mod_lm". As a metric I use the Root Mean Square Error (RMSE). it is a commonly used evaluation metric for regression tasks. RMSE measures the average deviation between the predicted values and the actual values, giving higher weight to larger errors. It provides a comprehensive understanding of the model's performance by considering both the magnitude and direction of errors.

```{r}
# Fit linear regression model
mod_lm <- caret::train(
  pp, 
  data = daily_fluxes_train |> drop_na(), 
  method = "lm",
  trControl = caret::trainControl(method = "none"),
  metric = "RMSE"
)

```


In this chunk I fit the KNN model "mod_knn". However, it is important to consider the number of neighbors. The selection of the number of neighbors in a k-NN model involves a trade-off between bias and variance. A smaller value of k, such as 1, can lead to a more flexible and low-bias model but may also make it more sensitive to noise and outliers. On the other hand, a larger value of k, such as 8, can help reduce the impact of noise and outliers, leading to a more robust and stable model with potentially lower variance.

```{r}
mod_knn <- caret::train(
  pp, 
  data = daily_fluxes_train |> drop_na(), 
  method = "knn",
  trControl = caret::trainControl(method = "none"),
  tuneGrid = data.frame(k = 8),
  metric = "RMSE"
)
```

The model objects mod_lm and mod_knn can be utilized to incorporate the fitted values into both the training and test data using the predict() function, along with the newdata argument. The following code snippet demonstrates the prediction step, the evaluation of prediction accuracy, and the visualization of predicted versus observed values on both the test and training sets. These functionalities are encapsulated within a single function named eval_model(), which can be reused for each fitted model object.

```{r}

# make model evaluation into a function to reuse code
eval_model <- function(mod, df_train, df_test){
  
  # add predictions to the data frames
  df_train <- df_train |> 
    drop_na()
  df_train$fitted <- predict(mod, newdata = df_train)
  
  df_test <- df_test |> 
    drop_na()
  df_test$fitted <- predict(mod, newdata = df_test)
  
  # get metrics tables
  metrics_train <- df_train |> 
    yardstick::metrics(GPP_NT_VUT_REF, fitted)
  
  metrics_test <- df_test |> 
    yardstick::metrics(GPP_NT_VUT_REF, fitted)
  
  # extract values from metrics tables
  rmse_train <- metrics_train |> 
    filter(.metric == "rmse") |> 
    pull(.estimate)
  rsq_train <- metrics_train |> 
    filter(.metric == "rsq") |> 
    pull(.estimate)
  
  rmse_test <- metrics_test |> 
    filter(.metric == "rmse") |> 
    pull(.estimate)
  rsq_test <- metrics_test |> 
    filter(.metric == "rsq") |> 
    pull(.estimate)
  
  # visualise as a scatterplot
  # adding information of metrics as sub-titles
  plot_1 <- ggplot(data = df_train, aes(GPP_NT_VUT_REF, fitted)) +
    geom_point(alpha = 0.3) +
    geom_smooth(method = "lm", se = FALSE, color = "red") +
    geom_abline(slope = 1, intercept = 0, linetype = "dotted") +
    labs(subtitle = bquote( italic(R)^2 == .(format(rsq_train, digits = 2)) ~~
                            RMSE == .(format(rmse_train, digits = 3))),
         title = "Training set") +
    theme_classic()
  
  plot_2 <- ggplot(data = df_test, aes(GPP_NT_VUT_REF, fitted)) +
    geom_point(alpha = 0.3) +
    geom_smooth(method = "lm", se = FALSE, color = "red") +
    geom_abline(slope = 1, intercept = 0, linetype = "dotted") +
    labs(subtitle = bquote( italic(R)^2 == .(format(rsq_test, digits = 2)) ~~
                            RMSE == .(format(rmse_test, digits = 3))),
         title = "Test set") +
    theme_classic()
  
  out <- cowplot::plot_grid(plot_1, plot_2)
  
  return(out)
}

# linear regression model
eval_model(mod = mod_lm, df_train = daily_fluxes_train, df_test = daily_fluxes_test)


```



```{r}
# KNN
eval_model(mod = mod_knn, df_train = daily_fluxes_train, df_test = daily_fluxes_test)


```


Interpretation of the results:
Larger Difference in Evaluation on Training and Test Set for KNN:
The larger difference between the evaluation on the training and test sets for the KNN model compared to the linear regression model suggests that the KNN model has higher variance. The KNN model tends to be more flexible and less biased, allowing it to closely fit the training data. However, this flexibility can also make it more susceptible to overfitting, resulting in a larger discrepancy between the model's performance on the training and test sets.

Better Model Performance on Test Set for KNN:
The evaluation on the test set indicating a better model performance for the KNN model compared to the linear regression model suggests that the KNN model has lower bias. The KNN model's ability to capture complex relationships and adapt to the data patterns allows it to generalize well to unseen data, resulting in improved performance on the test set. In contrast, the linear regression model may have a higher bias, meaning it makes stronger assumptions about the data and may not capture all the nuances present.

Positioning on the Bias-Variance Trade-Off Spectrum:
Considering the bias-variance trade-off, the KNN model is positioned towards the lower bias and higher variance end of the spectrum. It exhibits lower bias as it can flexibly capture complex relationships, but it has higher variance due to its sensitivity to the local structure of the training data. On the other hand, the linear regression model is positioned towards the higher bias and lower variance end. It makes stronger assumptions and has a simpler structure, resulting in higher bias but lower variance compared to the KNN model.

In a last step, I provide a visualization of the two models

Noch nicht alles visualisiert -> wie bringe ich test und training in einen plot?

```{r}

library(dplyr)
library(ggplot2)
library(hrbrthemes)


visualitation_test <- ggplot(daily_fluxes_test, aes(x=TIMESTAMP, y=GPP_NT_VUT_REF)) +
  geom_line( color="#69b3a2") + 
  xlab("") +
  theme_ipsum() +
  theme(axis.text.x=element_text(angle=60, hjust=1)) 
visualitation_test

visualitation_train <- ggplot(daily_fluxes_train, aes(x=TIMESTAMP, y=GPP_NT_VUT_REF)) +
  geom_line( color="#69b3a2") + 
  xlab("") +
  theme_ipsum() +
  theme(axis.text.x=element_text(angle=60, hjust=1)) 
visualitation_train


```



Exercice 2:

The In this 2nd subtask, the role of k is to be considered.
First question:
Hypothesis: As the number of neighbors (k) in a KNN model approaches 1, it is expected that the R2 (coefficient of determination) and MAE (Mean Absolute Error) evaluated on both the test and training sets would decrease. Conversely, as k approaches N (the number of observations in the data), it is expected that the R2 and MAE would increase. In summary, as k approaches 1, the KNN model becomes more overfit to the training data, leading to high training set performance but poor generalization. As k approaches N, the model becomes less flexible but more robust, resulting in lower training set performance but improved generalization to unseen data.


In the second question of this exercice I put the hypothesis to the test: 

To test the hypothesis, I use the data splitted into a training and test set above. For this I use a for loop to evaluate the variations in the different k's. To store the values of the k's, I define a list (list_values). The number of values is specified at 80. Such a high value was chosen to show the development of the Mae in detail.
In a first step, I make a model evaluation (eval_model2) into a function to reuse code.

```{r}

# make model evaluation into a function to reuse code
eval_model2 <- function(mod, df_train, df_test){
  
  # add predictions to the data frames
  df_train <- df_train |> 
    drop_na()
  df_train$fitted <- predict(mod, newdata = df_train)
  
  df_test <- df_test |> 
    drop_na()
  df_test$fitted <- predict(mod, newdata = df_test)
  
  # get metrics tables
  metrics_train <- df_train |> 
    yardstick::metrics(GPP_NT_VUT_REF, fitted)
  
  metrics_test <- df_test |> 
    yardstick::metrics(GPP_NT_VUT_REF, fitted)
  
  # extract values from metrics tables
  rmse_train <- metrics_train |> 
    filter(.metric == "rmse") |> 
    pull(.estimate)
  rsq_train <- metrics_train |> 
    filter(.metric == "rsq") |> 
    pull(.estimate)
  
  rmse_test <- metrics_test |> 
    filter(.metric == "rmse") |> 
    pull(.estimate)
  rsq_test <- metrics_test |> 
    filter(.metric == "rsq") |> 
    pull(.estimate)
  
  return(rmse_test)
}

# linear regression model
eval_model(mod = mod_lm, df_train = daily_fluxes_train, df_test = daily_fluxes_test)


```


```{r}

my_array <- seq(from = 1, to = 80)

list_values <- list()

library(tidyverse)
library(recipes)

for(k in 1:length(my_array)) {
  model_temp <- caret::train(
  pp, 
  data = daily_fluxes_train |> drop_na(), 
  method = "knn",
  trControl = caret::trainControl(method = "none"),
  tuneGrid = data.frame(k = k),
  metric = "MAE"
  )
  #Eval model
  list_values[k] <- eval_model2(model_temp, daily_fluxes_train, daily_fluxes_test)
}


```

It appears that the optimal value of k in the k-nearest neighbors (KNN) model is determined to be 26, as it yields the smallest error with a value of 1.53. This result suggests that when considering the model complexity, a moderate level of nearest neighbors yields the best performance in terms of minimizing the Mean Absolute Error (MAE) metric. The fact that the MAE reaches its minimum at k = 26 indicates that increasing the complexity of the model up to this point leads to improved predictions on the unseen test set. However, the larger the k the better the prediction in the training dataset and the worse the flexibility when I apply the model to data that is independent of the original data.



```{r}

# Create dataframe from list_values
results_df <- data.frame(k = my_array, metric_value = unlist(list_values))

# Plotting
ggplot(results_df, aes(x = k, y = metric_value)) +
  geom_line() +
  geom_point() +
  labs(x = "Model Complexity (k)", y = "Evaluation Metric") +
  ggtitle("Model Generalizability as a Function of Model Complexity")



```

In the visual representation of model generalizability as a function of model complexity,  the regions of overfitting and underfitting can be identified by observing the behavior of the evaluation metric (e.g., MAE) as the model complexity increases.
Overfitting region: In the case of overfitting, the model complexity is high, which means the model is capturing noise or random variations in the training data rather than the underlying patterns. In the plot, you may notice that as the model complexity increases, the evaluation metric on the training set decreases (e.g., MAE decreases). However, at the same time, the evaluation metric on the test/validation set may start to increase or plateau, or the rate of improvement becomes slower. This indicates that the model is becoming too specific to the training data and is failing to generalize well to new, unseen data. The overfitting region is characterized by a large gap between the evaluation metric values on the training set and the test/validation set.
Underfitting region: On the other hand, in the case of underfitting, the model complexity is too low to capture the underlying patterns in the data. In the plot, you may observe that the evaluation metric on both the training set and the test/validation set remains relatively high and does not show significant improvement as the model complexity increases. This indicates that the model is not able to adequately learn from the training data and is failing to capture the underlying relationships or patterns. The underfitting region is characterized by high evaluation metric values on both the training set and the test/validation set.








